---
title: "known_fusions"
output: html_notebook
---


##KNOWN ABCB1 FUSIONS - VALIDATION
-paper:
i.GBM	TCGA-19-2620-01A-01R-1850-01	GTF2IP4--ABCB1
ii.HNSC	TCGA-BB-7872-01A-11R-2232-07	TP53TG1--ABCB1
iii.LIHC	TCGA-BC-A10T-01A-11R-A131-07	XRCC5--ABCB1
iv.LIHC	TCGA-DD-A73F-01A-11R-A32O-07	CAPZA2--ABCB1
-tumorfusions
LIHC	TCGA.DD.A73F.01A	CAPZA2::ABCB1	
LIHC	TCGA.BC.A10T.01A	XRCC5::ABCB1

```{r}
tcga_normalised_df

#i - rail_69514
subset(colData(abcb1.exon), 
       gdc_cases.samples.portions.analytes.aliquots.submitter_id == "TCGA-19-2620-01A-01R-1850-01", 
       select = c("rail_id", "gdc_cases.samples.submitter_id", "cgc_case_histological_diagnosis", "cgc_drug_therapy_drug_name"))
#ii - not present
subset(colData(abcb1.exon), 
       gdc_cases.samples.portions.analytes.aliquots.submitter_id == "TCGA-BB-7872-01A-11R-2232-07", 
       select = c("rail_id", "gdc_cases.samples.submitter_id", "cgc_case_histological_diagnosis", "cgc_drug_therapy_drug_name"))
#iii - rail_61860
subset(colData(abcb1.exon), 
       gdc_cases.samples.portions.analytes.aliquots.submitter_id == "TCGA-BC-A10T-01A-11R-A131-07", 
       select = c("rail_id", "gdc_cases.samples.submitter_id", "cgc_case_histological_diagnosis", "cgc_drug_therapy_drug_name"))
#iv - rail_60306
subset(colData(abcb1.exon), 
       gdc_cases.samples.portions.analytes.aliquots.submitter_id == "TCGA-DD-A73F-01A-11R-A32O-07", 
       select = c("rail_id", "gdc_cases.samples.submitter_id", "cgc_case_histological_diagnosis", "cgc_drug_therapy_drug_name"))

known.fusions <- tcga_normalised_df[c("rail_69514", "rail_61860", "rail_60306"),]
known.fusions.liver <- known.fusions[2:3,]
known.fusions
known.fusions.liver
```

```{r}
known.fusions.liver_long <- known.fusions.liver[-c(30:40)]
known.fusions.liver_long <- data.frame(t(known.fusions.liver_long))
known.fusions.liver_long$exon_number <- c(1:29)

tcga_liver_df <- subset(tcga_normalised_df, tissue == "Liver")
tcga_liver_df
gtex_liver_df <- subset(gtex_normalised_df, tissue == "Liver")
gtex_liver_df

tcga_liver_long <- tcga_liver_df[-c(30:40)]
tcga_liver_long <- data.frame(t(tcga_liver_long))
tcga_liver_long$exon_number <- c(1:29)

gtex_liver_long <- gtex_liver_df[-c(30:47)]
gtex_liver_long <- data.frame(t(gtex_liver_long))
gtex_liver_long$exon_number <- c(1:29)
```





```{r}
known.fusions.liver_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-15:2), limits = c(-16,3))

tcga_liver_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-15:2), limits = c(-16,3)) +
  labs(title = "TCGA Liver data")

gtex_liver_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-15:2), limits = c(-16,3)) +
  labs(title = "GTEx Liver data")
```

```{r}

graph.known <- melt(known.fusions.liver_long, id.vars = "exon_number")
graph.liver <- melt(tcga_liver_long, id.vars = "exon_number")

ggplot(graph.liver, aes(x = exon_number, y = log2(value), group = variable)) +
  geom_line() +
  geom_point() +
  geom_line(data = graph.known, colour = "red") +
  geom_point(data = graph.known, colour = "red") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-15:2), limits = c(-16,3)) 

ggplot(graph.liver, aes(x = exon_number, y = value, group = variable)) +
  geom_line() +
  geom_point() +
  geom_line(data = graph.known, colour = "red") +
  geom_point(data = graph.known, colour = "red") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) 
```


```{r}
rbind(data.frame(t(known.fusions.liver[, c("tcga_id", "one", "total_counts", "tissue", "tumour_type", "treatment")])),
data.frame(t(data.frame(colData(abcb1.exon)[c("rail_61860", "rail_60306"), c("gdc_file_id", "gdc_cases.samples.sample_type", "gdc_cases.demographic.year_of_birth", "cgc_case_year_of_diagnosis", "cgc_case_tumor_status", "gdc_cases.demographic.year_of_death", "cgc_slide_percent_tumor_nuclei", "cgc_slide_percent_normal_cells", "cgc_slide_percent_tumor_cells", "cgc_slide_percent_necrosis")]))))
```
```{r}
tcga_liver_df
```

```{r}
tcga_liver_mean_long <- data.frame(rowMeans(tcga_liver_long[,-425]))
tcga_liver_mean_long$exon_number <- c(1:29)
tcga_liver_mean_long <- melt(tcga_liver_mean_long, id.vars = "exon_number")
tcga_liver_mean_long

tcga_liver_mean_long %>%
  ggplot(aes(x = exon_number, y = value)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none")

gtex_liver_mean_long <- data.frame(rowMeans(gtex_liver_long[,-137]))
gtex_liver_mean_long$exon_number <- c(1:29)
gtex_liver_mean_long <- melt(gtex_liver_mean_long, id.vars = "exon_number")

gtex_liver_mean_long %>%
  ggplot(aes(x = exon_number, y = value)) +
  geom_line() +
  geom_point()


ggplot(NULL, aes(exon_number, log2(value), group = variable, colour = variable)) +
  geom_line(data = gtex_liver_mean_long, size = 2) +
  geom_point(data = gtex_liver_mean_long, size = 3) +
  geom_line(data = tcga_liver_mean_long, size = 2) +
  geom_point(data = tcga_liver_mean_long, size = 3) +
  geom_line(data = melt(known.fusions.liver_long, id.vars = "exon_number")) +
  geom_point(data = melt(known.fusions.liver_long, id.vars = "exon_number")) +
  labs(title = "ABCB1 data: TCGA Liver fusions, TCGA & GTEx mean") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-14:3), limits = c(-14,3))
```

#potential fusions vs known fusions

```{r}
ggplot(NULL, aes(exon_number, log2(value), group = variable, colour = variable)) +
  geom_line(data = melt(tcga_ovary_long[,c(1:3,430)], id.vars = "exon_number")) +
  geom_point(data = melt(tcga_ovary_long[,c(1:3,430)], id.vars = "exon_number")) +
  geom_line(data = melt(known.fusions.liver_long, id.vars = "exon_number"), size = 2) +
  geom_point(data = melt(known.fusions.liver_long, id.vars = "exon_number"), size = 3) +
  labs(title = "ABCB1 data: known fusions vs potential fusions") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-14:3), limits = c(-14,3))

ggplot(NULL, aes(exon_number, value, group = variable, colour = variable)) +
  geom_line(data = melt(tcga_ovary_long[,c(1:3,430)], id.vars = "exon_number")) +
  geom_point(data = melt(tcga_ovary_long[,c(1:3,430)], id.vars = "exon_number")) +
  geom_line(data = melt(known.fusions.liver_long, id.vars = "exon_number"), size = 2) +
  geom_point(data = melt(known.fusions.liver_long, id.vars = "exon_number"), size = 3) +
  labs(title = "ABCB1 data: known fusions vs potential fusions") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) 
```

```{r}
tcga_liver_long[420:424]
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = value, group = variable, colour = variable)) +
  geom_point() +
  geom_line() +
  theme(legend.position = "none")

tcga_liver_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_point() +
  geom_line() +
  theme(legend.position = "none")
```

```{r}
summary(tcga_liver_df$total_counts)
tcga_liver_df[c("rail_60306", "rail_61860"),]
```



ok want to compare 1-19 to 20-29
yo so this other known fusion is just to exon 29 - explains why that's more highly expressed - more obvious in raw
```{r}
known.fusions.liver$fusion_ratio <- (rowSums(known.fusions.liver[,20:29])/10)/(rowSums(known.fusions.liver[,1:19])/19)
known.fusions.liver

tcga_liver_df$fusion_ratio <- (rowSums(tcga_liver_df[,20:29])/10)/(rowSums(tcga_liver_df[,1:19])/19)
tcga_liver_df

summary(known.fusions.liver$fusion_ratio)
summary(tcga_liver_df$fusion_ratio) #wait hold up are the samples not in with the rest of the liver?

#we normalise stuff with log2
wilcox.test(known.fusions.liver$fusion_ratio, tcga_liver_df$fusion_ratio)

```

```{r}
#tcga_liver_df[-c("rail_61660", "rail_70512", "rail_64067"),]

cor.test(tcga_liver_df$fusion_ratio, tcga_liver_df$total_counts, method = "kendall")
cor.test(tcga_liver_df$fusion_ratio, tcga_liver_df$total_counts, method = "spearman")

cor.test(liver_without_outliers$fusion_ratio, liver_without_outliers$total_counts, method = "kendall")
cor.test(liver_without_outliers$fusion_ratio, liver_without_outliers$total_counts, method = "spearman")
```

```{r}
tcga_liver_df %>%
  ggplot(aes(x = total_counts, y = fusion_ratio)) +
  geom_point() +
  theme(legend.position = "none") +
  scale_y_continuous(limits = c(0,25))
```
```{r}
summary(tcga_liver_df$total_counts)
tcga_liver_df[order(tcga_liver_df$fusion_ratio, decreasing = TRUE),]
```



```{r}
tcga_raw_df[c("rail_61860", "rail_60306"),]
```






