---
title: "Stats"
output: html_notebook
---

##time for some stats
##can you do a t test for one sample???? - because that won't be a mean
##want to compare exon data, do all the tests for exon 1, exon 2, and maybe for my og definition of exon 1
let's just do lots of tests and find some stuff
##ooh also just total exon count so we can get a thing for overexpression
-potential fusions and known fusions
-known fusions and gtex
-potential fusions and gtex

Let's do t tests
-explanatory (exon)
-response (mean number of counts - across all the samples)
-see this doesn't feel like two separate variables

-ooh wait - explanatory is the kind of data: normal, cancer, potential fusion, known fusion
-response is still number of counts
-think it's ok that the two groups have different amounts 

t tests require the groups to be normally distributed, and variances to be =

"The test statistic evaluates how well the result observed matches the
result expected if the null is true"

R struc: t.test(response, explanatory) - not sure if t test is what I want then - doesn't exactly follow the response, explanatory format
-i think the one  i'm mpst interested in is the response: cancer, potential etc
#Potential fusions and known fusions
#this test we actually don't want significant results
-sum of all exons
H0: sum(potential) = sum(known)
H1: sum(potential) /= sum(known)
-exon 1
H0: sum(potential) = sum(known)
H1: sum(potential) /= sum(known)
-exon 2
H0: sum(potential) = sum(known)
H1: sum(potential) /= sum(known)

I think I want z test for the gtex and ovary - lots of samples, and t test for my known and potential?

```{r}
#First - are the groups normally distributed? Shapiro-Wilk test
if(!require(devtools)) install.packages("devtools")
devtools::install_github("kassambara/ggpubr")
library("ggpubr")

```
"The central limit theorem tells us that no matter what distribution things have, the sampling distribution tends to be normal if the sample is large enough (n > 30)."
-well that's a point in my favour

```{r}
tcga_ovary_df$log2_total_counts <- mutate(tcga_ovary_df, log2_total_counts = log2(total_counts))
gtex_ovary_df$log2_total_counts <- mutate(gtex_ovary_df, log2_total_counts = log2(total_counts))
ggdensity(tcga_ovary_df$log2_total_counts) 

ggqqplot(tcga_ovary_df$log2_total_counts)
t.test(tcga_ovary_df$log2_total_counts, gtex_ovary_df$log2_total_counts, var.equal = FALSE)
t.test(tcga_ovary_df$total_counts, gtex_ovary_df$total_counts, var.equal = FALSE)
```

```{r}
ggplot(tcga_ovary_df, aes(x = log2(total_counts))) +
  geom_histogram()

ggplot(gtex_ovary_df, aes(x = log2(total_counts))) +
  geom_histogram()

#shapiro.test(tcga_ovary_df$log2_total_counts)

treatment.na_df <- tcga_ovary_df[is.na(tcga_ovary_df$treatment),]
treated_df <- tcga_ovary_df[!is.na(tcga_ovary_df$treatment),]

#difference between treated and untreated is not statistically significant
t.test(treated_df$total_counts, treatment.na_df$total_counts, var.equal = FALSE)
wilcox.test(treated_df$total_counts, treatment.na_df$total_counts)


#Rosner's test: can detect multiple outliers at once, best for large datasets

#library(EnvStats)
rosner <- rosnerTest(tcga_ovary_df$total_counts, k = 10)
rosner$all.stats

rosner <- rosnerTest(gtex_ovary_df$total_counts, k = 10)
rosner$all.stats

#so thought this test was good but actually it says the top 6 results are outliers - I think it's the log transformation that really shows the main ones
```
```{r}
tcga_ovary_df$log2_total_counts = log2(tcga_ovary_df$total_counts)
tcga_ovary_df
gtex_ovary_df$log2_total_counts = log2(gtex_ovary_df$total_counts)
gtex_ovary_df

t.test(tcga_ovary_df$log2_total_counts, gtex_ovary_df$log2_total_counts, var.equal = FALSE)

treatment.na_df <- tcga_ovary_df[is.na(tcga_ovary_df$treatment),]
treated_df <- tcga_ovary_df[!is.na(tcga_ovary_df$treatment),]

t.test(treated_df$log2_total_counts, treatment.na_df$log2_total_counts, var.equal = FALSE)
wilcox.test(treated_df$total_counts, treatment.na_df$total_counts)
#oh here's a thought
#exon 1 compared to average for the rest
#I really do just need to stop
```

```{r}
summary(tcga_ovary_df$one)
summary(tcga_ovary_df$three)

summary(gtex_ovary_df$one)
summary(gtex_ovary_df$three)
```

so I could do a z test - for exon 1 (expressed) to see how it's different from the complete average of expression
```{r}
stats_tcga_ovary_df <- tcga_ovary_df
#rowMeans(stats_tcga_ovary_df[,-c(30:41)])
colMeans(stats_tcga_ovary_df[,-c(30:41)])
mean(colMeans(stats_tcga_ovary_df[,-c(30:41)]))
mean(colMeans(filter(stats_tcga_ovary_df, one > 0)[,-c(30:41)]))
mean(colMeans(filter(stats_tcga_ovary_df, one == 0)[,-c(30:41)]))

stats_gtex_ovary_df <- gtex_ovary_df
#rowMeans(stats_gtex_ovary_df[,-c(30:45)])
colMeans(stats_gtex_ovary_df[,-c(30:45)])
mean(colMeans(stats_gtex_ovary_df[,-c(30:45)]))
mean(colMeans(filter(stats_gtex_ovary_df, one > 0)[,-c(30:45)]))
mean(colMeans(filter(stats_gtex_ovary_df, one == 0)[,-c(30:45)]))
```
so that was getting complicated (and confusing) - what about looking at above a certain threshold - I know that's bad but I just want to look at the interesting stuff when it's actually being expressed

```{r}
summary(stats_tcga_ovary_df$total_counts)
summary(stats_gtex_ovary_df$total_counts)

summary(stats_tcga_ovary_df$log2_total_counts)
summary(stats_gtex_ovary_df$log2_total_counts)
```

```{r}
mean(colMeans(filter(stats_tcga_ovary_df, total_counts > 0.290841)[,-c(30:41)]))
sd(colMeans(filter(stats_tcga_ovary_df, total_counts > 0.290841)[,-c(30:41)]))

colMeans(filter(stats_tcga_ovary_df, total_counts > 0.290841)[,-c(30:41)])
```


```{r}
#stats_tcga_ovary_df$quartile <- ntile(stats_tcga_ovary_df$total_counts, 4)
#kruskal.test(total_counts ~ quartile, data = stats_tcga_ovary_df)
#pairwise.wilcox.test(stats_tcga_ovary_df$total_counts, stats_tcga_ovary_df$quartile,
                   #  p.adjust.method = "BH")
kruskal.test(one ~ quartile, data = stats_tcga_ovary_df)
pairwise.wilcox.test(stats_tcga_ovary_df$one, stats_tcga_ovary_df$quartile,
                     p.adjust.method = "BH") #tada - this is very good
```

```{r}
#stats_gtex_ovary_df$quartile <- ntile(stats_gtex_ovary_df$total_counts, 4)
#kruskal.test(total_counts ~ quartile, data = stats_gtex_ovary_df)
#pairwise.wilcox.test(stats_gtex_ovary_df$total_counts, stats_gtex_ovary_df$quartile,
        #             p.adjust.method = "BH")
kruskal.test(one ~ quartile, data = stats_gtex_ovary_df)
pairwise.wilcox.test(stats_gtex_ovary_df$one, stats_gtex_ovary_df$quartile,
                     p.adjust.method = "BH")
```

```{r}
pairwise.wilcox.test(stats_tcga_ovary_df$one, stats_tcga_ovary_df$quartile,
                     p.adjust.method = "BH")
pairwise.wilcox.test(stats_gtex_ovary_df$one, stats_gtex_ovary_df$quartile,
                     p.adjust.method = "BH")

```

```{r}
stats_tcga_ovary_df %>%
  filter(quartile == 4) %>% #107
  filter(one == 0) #41
41/107 #38%

stats_gtex_ovary_df %>%
  filter(quartile == 4) %>% #27
  filter(one == 0) #4
4/27 #15%

treated_df
treatment.na_df
```


```{r}
summary(filter(stats_tcga_ovary_df, quartile == 4)$one)
summary(filter(stats_tcga_ovary_df, quartile == 3)$one)
summary(filter(stats_tcga_ovary_df, quartile == 2)$one)
summary(filter(stats_tcga_ovary_df, quartile == 1)$one)
wilcox.test(filter(stats_tcga_ovary_df, quartile == 4)$one, filter(stats_tcga_ovary_df, quartile == 3)$one) #nice that worked

stats_tcga_ovary_df$quartile <- as.factor(stats_tcga_ovary_df$quartile)
stats_tcga_ovary_df %>%
  ggplot(aes(x = quartile, one)) +
  geom_boxplot()

stats_tcga_ovary_df %>%
  ggplot(aes(x = quartile, total_counts)) +
  geom_boxplot()

stats_gtex_ovary_df$quartile <- as.factor(stats_gtex_ovary_df$quartile)
stats_gtex_ovary_df %>%
  ggplot(aes(x = quartile, one)) +
  geom_boxplot()

stats_gtex_ovary_df %>%
  ggplot(aes(x = quartile, total_counts)) +
  geom_boxplot()


stats_tcga_ovary_df %>%
  ggplot(aes(x = quartile, log2(one + 1))) +
  geom_boxplot()

stats_gtex_ovary_df %>%
  ggplot(aes(x = quartile, log2(one + 1))) +
  geom_boxplot()
```
```{r}
stats_tcga_ovary_df %>%
  ggplot(aes(x = quartile, y = total_counts)) +
  geom_boxplot() +
  geom_boxplot(y = one, colour = "red") +
  scale_y_continuous("Total Counts", sec.axis = sec_axis(name = "One"))
```


maybe can combine one and total by quantile using facet or something
would be nice to have all 4 in 1 figure

also let's do a correlation because I'm on a roll
non-normal: Kendall and Spearman
```{r}
cor.test(stats_tcga_ovary_df$total_counts, stats_tcga_ovary_df$one, method = "kendall")
cor.test(stats_tcga_ovary_df$total_counts, stats_tcga_ovary_df$one, method = "spearman")

cor.test(stats_gtex_ovary_df$total_counts, stats_gtex_ovary_df$one, method = "kendall")
cor.test(stats_gtex_ovary_df$total_counts, stats_gtex_ovary_df$one, method = "spearman")
```

```{r}
cor.test(filter(stats_tcga_ovary_df, quartile == 4)$total_counts, filter(stats_tcga_ovary_df, quartile == 4)$one, method = "kendall")
cor.test(filter(stats_tcga_ovary_df, quartile == 4)$total_counts, filter(stats_tcga_ovary_df, quartile == 4)$one, method = "spearman")

cor.test(filter(stats_gtex_ovary_df, quartile == 4)$total_counts, filter(stats_gtex_ovary_df, quartile == 4)$one, method = "kendall")
cor.test(filter(stats_gtex_ovary_df, quartile == 4)$total_counts, filter(stats_gtex_ovary_df, quartile == 4)$one, method = "spearman")
```

```{r}
stats_tcga_ovary_df %>%
  ggplot(aes(x = log2(total_counts + 1), y = log2(one + 1), colour = quartile)) +
  geom_point()

stats_gtex_ovary_df %>%
  ggplot(aes(x = total_counts, y = one, colour = quartile)) +
  geom_point()

ggplot(NULL, aes(x = log2(total_counts + 1), y = log2(one + 1), colour = quartile)) +
  geom_point(data = stats_tcga_ovary_df) +
  geom_point(data = stats_gtex_ovary_df, shape = "triangle")
```

```{r}
stats_tcga_ovary_df
```
```{r}
summary(tcga_ovary_df$total_counts)
```


```{r}
stats_tcga_ovary_df %>%
  ggplot(aes(x = quartile, one)) +
  geom_boxplot()

stats_tcga_ovary_df %>%
  ggplot(aes(x = quartile, total_counts)) +
  geom_boxplot() 

stats_gtex_ovary_df %>%
  ggplot(aes(x = quartile, one)) +
  geom_boxplot()

stats_gtex_ovary_df %>%
  ggplot(aes(x = quartile, total_counts)) +
  geom_boxplot()
```

```{r}
stats_gtex_ovary_df$source <- "GTEX"
stats_tcga_ovary_df$source <- "TCGA"
```



```{r}
stats_tcga_ovary_df
stats_gtex_ovary_df
```


```{r}
rbind(stats_tcga_ovary_df[,-c(30,32:33,35:40)], stats_gtex_ovary_df[,-c(30,32:34,36:44)])
```



```{r}
summary(filter(stats_gtex_ovary_df, quartile == 4)$one)
summary(filter(stats_gtex_ovary_df, quartile == 3)$one)
summary(filter(stats_gtex_ovary_df, quartile == 2)$one)
summary(filter(stats_gtex_ovary_df, quartile == 1)$one)
wilcox.test(filter(stats_gtex_ovary_df, quartile == 4)$one, filter(stats_gtex_ovary_df, quartile == 3)$one) #well that gave different answer to pairwise so...
```



```{r}
wilcox.test(tcga_cervix_df$total_counts, gtex_cervix_df$total_counts)
wilcox.test(tcga_kidney_df$total_counts, gtex_kidney_df$total_counts)
wilcox.test(tcga_bladder_df$total_counts, gtex_bladder_df$total_counts)
wilcox.test(tcga_stomach_df$total_counts, gtex_stomach_df$total_counts)
wilcox.test(tcga_breast_df$total_counts, gtex_breast_df$total_counts)
wilcox.test(tcga_liver_df$total_counts, gtex_liver_df$total_counts)
wilcox.test(tcga_lung_df$total_counts, gtex_lung_df$total_counts)
wilcox.test(tcga_testis_df$total_counts, gtex_testis_df$total_counts)
wilcox.test(tcga_blood_df$total_counts, gtex_blood_df$total_counts)
wilcox.test(tcga_brain_df$total_counts, gtex_brain_df$total_counts)
wilcox.test(tcga_colon_df$total_counts, gtex_colon_df$total_counts)
```
```{r}
wilcox.test(tcga_lung_df$total_counts, gtex_lung_df$total_counts, alternative = "less")
wilcox.test(tcga_lung_df$total_counts, gtex_lung_df$total_counts, alternative = "less")
wilcox.test(tcga_testis_df$total_counts, gtex_testis_df$total_counts, alternative = "less")
wilcox.test(tcga_blood_df$total_counts, gtex_blood_df$total_counts, alternative = "less") #nope p = 1
wilcox.test(tcga_breast_df$total_counts, gtex_breast_df$total_counts, alternative = "less")
wilcox.test(tcga_bladder_df$total_counts, gtex_bladder_df$total_counts, alternative = "less")
wilcox.test(tcga_cervix_df$total_counts, gtex_cervix_df$total_counts, alternative = "less")
```
```{r}
tcga_colon_df
gtex_colon_df
```


```{r}
wilcox.test(tcga_brain_df$total_counts, gtex_brain_df$total_counts, alternative = "greater")
wilcox.test(tcga_liver_df$total_counts, gtex_liver_df$total_counts, alternative = "greater")
wilcox.test(tcga_colon_df$total_counts, gtex_colon_df$total_counts, alternative = "greater")
wilcox.test(tcga_kidney_df$total_counts, gtex_kidney_df$total_counts, alternative = "greater")
wilcox.test(tcga_stomach_df$total_counts, gtex_stomach_df$total_counts, alternative = "greater") #just over p = 0.05893
```

```{r}
select_tcga <- thesis_graph_combined %>%
  filter(tissue != "NA") %>%
  filter(tissue != "Adipose Tissue") %>%
  filter(tissue != "Bile Duct") %>%
  filter(tissue != "Blood Vessel") %>%
  filter(tissue != "Bone Marrow") %>%
  filter(tissue != "Eye") %>%
  filter(tissue != "Head and Neck") %>%
  filter(tissue != "Heart") %>%
  filter(tissue != "Lymph Nodes") %>%
  filter(tissue != "Mesenchymal") %>%
  filter(tissue != "Mesothelium") %>%
  filter(tissue != "Muscle") %>%
  filter(tissue != "Salivary Gland") %>%
  filter(tissue != "Spleen") %>%
  filter(tissue != "Thymus") %>%
  filter(tissue != "") %>%
  filter(tissue != "Vagina") %>%
  filter(tissue != "Adrenal Gland") %>%
  filter(tissue != "Nerve") %>%
  filter(tissue != "Nervous System") %>%
  filter(tissue != "Pituitary") %>%
  filter(tissue != "Blood") %>%
  filter(tissue != "Small Intestine") %>%
  filter(tissue != "Thyroid") %>%
  filter(tissue != "Fallopian Tube") %>%
  filter(tissue != "Skin") %>%
  filter(tissue != "Esophagus") %>% #where it was before decided to reduce even further
  filter(tissue != "Pancreas") %>%
  filter(tissue != "Prostate") %>%
  filter(tissue != "Uterus") %>%
  filter(source == "TCGA")
```



```{r}
wilcox.test(filter(select_tcga, tissue == "Cervix")$total_counts ~ filter(select_tcga, tissue == "Cervix")$treated)
wilcox.test(filter(select_tcga, tissue == "Colon/Rectal")$total_counts ~ filter(select_tcga, tissue == "Colon/Rectal")$treated)
wilcox.test(filter(select_tcga, tissue == "Breast")$total_counts ~ filter(select_tcga, tissue == "Breast")$treated)
wilcox.test(filter(select_tcga, tissue == "Stomach")$total_counts ~ filter(select_tcga, tissue == "Stomach")$treated)
wilcox.test(filter(select_tcga, tissue == "Brain")$total_counts ~ filter(select_tcga, tissue == "Brain")$treated)
wilcox.test(filter(select_tcga, tissue == "Bladder")$total_counts ~ filter(select_tcga, tissue == "Bladder")$treated)
wilcox.test(filter(select_tcga, tissue == "Kidney")$total_counts ~ filter(select_tcga, tissue == "Kidney")$treated)
wilcox.test(filter(select_tcga, tissue == "Liver")$total_counts ~ filter(select_tcga, tissue == "Liver")$treated)
wilcox.test(filter(select_tcga, tissue == "Lung")$total_counts ~ filter(select_tcga, tissue == "Lung")$treated)
wilcox.test(filter(select_tcga, tissue == "Stomach")$total_counts ~ filter(select_tcga, tissue == "Stomach")$treated)
wilcox.test(filter(select_tcga, tissue == "Testis")$total_counts ~ filter(select_tcga, tissue == "Testis")$treated)
```


```{r}
pairwise.wilcox.test(filter(thesis_graph_combined, source == "TCGA")$total_counts, filter(thesis_graph_combined, source == "TCGA")$tissue,
                    p.adjust.method = "BH")


```


```{r}
ggplot(tcga_ovary_df, aes(x = total_counts)) +
  geom_histogram()
ggplot(tcga_ovary_df, aes(x = log2(total_counts + 1))) +
  geom_histogram()

ggplot(gtex_ovary_df, aes(x = total_counts)) +
  geom_histogram()
ggplot(gtex_ovary_df, aes(x = log2(total_counts + 1))) +
  geom_histogram()

t.test(mutate(tcga_ovary_df, log2_total_counts = log2(total_counts))$log2_total_counts, mutate(gtex_ovary_df, log2_total_counts = log2(total_counts))$log2_total_counts, var.equal = FALSE)
t.test(mutate(tcga_ovary_df, log2_total_counts = log2(total_counts + 1))$log2_total_counts, mutate(gtex_ovary_df, log2_total_counts = log2(total_counts + 1))$log2_total_counts, var.equal = FALSE)
```
```{r}
combined_df %>%
  filter(source = "TCGA") %>%
  
```




-normality check for total_counts data
```{r}
excel_known.df$total_counts <- rowSums(excel_known.df)
excel_known.df
#visually check normality
#density plot - want bell shape - definitely not bell shape
ggdensity(excel_known.df$total_counts)
#qq plot - correlation between sample and ditrib - wants points to fall on line - they're actually close
ggqqplot(excel_known.df$total_counts)

#visual checks aren't that great - let's do analysis
shapiro.test(excel_known.df$total_counts)
#get p = 0.3679 > 0.05 - assume normality
#let's do a variance test: known and potential fusions

known.sum <- c(excel_known.df$total_counts)
potential.sum <- c(potentialbb.df$total_counts)
tcga.sum <- c(tcga_ovary_df$total_counts)
gtex.sum <- c(gtex_ovary_norm_df$total_counts)
var.test(known.sum, potential.sum, var.equal = TRUE) #p=0.008 <0.05 - variances not equal
```


```{r}
tcga_ovary_df
ggdensity(tcga_ovary_df[-c(1:3),]$total_counts) #definitely not bell
ggqqplot(tcga_ovary_df$total_counts) #most on line but some massive outliers
ggqqplot(tcga_ovary_df[-c(1:3),]$total_counts)
shapiro.test(tcga_ovary_df[-c(1:3),]$total_counts) #p very very close to 0 < 0.05 - not normal data  - could be because of large sample size - anything becomes significant
```



```{r}
potentialbb.df <- tcga_ovary_df[c("rail_60851", "rail_63388", "rail_62767", "rail_68551", "rail_64510"),]
potentialbb.df
ggdensity(potentialbb.df$total_counts) #wave like known
ggqqplot(potentialbb.df$total_counts) #pretty close to line
shapiro.test(potentialbb.df$total_counts) #p=0.15 > 0.05 - normal

gtex_ovary_norm_df$total_counts <- rowSums(gtex_ovary_norm_df[1:29])
gtex_ovary_norm_df
ggdensity(gtex_ovary_norm_df$total_counts) #skewed bell
ggqqplot(gtex_ovary_norm_df$total_counts) #fairly close to line
shapiro.test(gtex_ovary_norm_df$total_counts) # p < 0.05 - probably because lots of samples - anything significant then

#make some vectors
known.one <- c(excel_known.df$one)
potential.one <- c(potentialbb.df$one)
tcga.one <- c(tcga_ovary_df$one)
gtex.one <- c(gtex_ovary_norm_df$one)

known.two <- c(excel_known.df$two)
potential.two <- c(potentialbb.df$two)
tcga.two <- c(tcga_ovary_df$two)
gtex.two <- c(gtex_ovary_norm_df$two)

var.test(known.one, potential.one, var.equal = TRUE) #var equal
var.test(known.two, potential.two, var.equal = TRUE)
```

let's to a t test - well the non-parametric one because it failed that
-sum: potential and known
```{r}
t.test(known.sum, potential.sum, var.equal = FALSE) #p=0.28
```
p = 0.28 > 0.05 - not significant 
this is good, this is what we wanted
means difference in means is not significant

t.test(x,y)
t.test(response, explanatory)
x-y=0
x-y>0 - then x>y 

difference in means not 0. greater: difference in means >0, less: difference in means <0
```{r}
t.test(known.sum, gtex.sum, var.equal = FALSE, alternative = "greater") #p=0.21 greater = 0.1, less=0.89
```

```{r}
t.test(potential.sum, gtex.sum, var.equal = FALSE, alternative = "greater") #p=0.06, greater =0.03 I wonder if it would be significant if I just used the 2 main ones?

t.test(potential.sum, tcga.sum, var.equal = FALSE, alternative = "greater") # p=0.03, greater = 0.019- good, the potential fusions are signif different from tcga
```
-this is good
signif potential > gtex (0.03)
signif potential > tcga (0.019)
so more signif with tcga
tcga < gtex < potential
-can officially say yes, these samples are significantly overexpressed 

```{r}
t.test(tcga.sum, gtex.sum, var.equal = FALSE, alternative = "less") #p=2.2e-16, greater=1, less=2.2e-16
```

-this says gtex sum > tcga sum
-I guess this makes sense in the whole lots of mutations and looks like lots of exon skipping in tcga - so overall - sum is less



#nothing significant for comparing exon 1
-most importantly - there's no significant difference in mean for exon 1 between gtex and tcga, and potential - though potential was close 0.11 - might be significant if I only use the 2 most important samples?
```{r}
t.test(known.one, potential.one, var.equal = FALSE) #p=0.4 - good not signif
t.test(known.one, gtex.one, var.equal = FALSE, alternative = "less") #p=0.5 - sad not signif - closer when alt = greater p=0.2, alt=less p=0.7
t.test(known.one, tcga.one, var.equal = FALSE, alternative = "less") #p=0.5 - not good, alt = greater p=0.2, less=0.7
t.test(tcga.one, gtex.one, var.equal = FALSE, alternative = "less") #p=0.9, alt greater=0.54, less = 0.45
t.test(potential.one, gtex.one, var.equal = FALSE, alternative = "greater") #p=0.2, greater=0.1134, less=0.88
t.test(potential.one, tcga.one, var.equal = FALSE, alternative = "less") #p=0.2, greater =0.11, less=0.88
```


#interesting stuff for exon 2
gtex - tcga >0 significant  - meaning: tcga exon 2 < gtex exon 2 on average
this - asides from the total sum - is the significant difference between the normal and cancer data
so focus significance wise should be on non-coding exon 2 - as long as what I learn from Liz doesn't contradict it

potential - gtex >0 signfic : gtex < potential
potential - tcga >0 signif: tcga < potential  - but more signif than (potential - gtex) >0

so I could order it:   tcga < gtex < potential which feels odd but we'll see
I think I need to do the 2 known liver cancer ones against liver cancer and liver gtex to properly compare
```{r}
t.test(known.two, potential.two, var.equal = FALSE) #p=0.8 - good
t.test(known.two, gtex.two, var.equal = FALSE, alternative = "less") #p=0.2, greater =0.11, less =0.8
t.test(known.two, tcga.two, var.equal = FALSE, alternative = "less") # p=0.2, greater =0.11, less=0.89
t.test(tcga.two, gtex.two, var.equal = FALSE, alternative = "less") #p=0.0001 - signif, greater=0.999, less=9.59e-05#interesting
t.test(potential.two, gtex.two, var.equal = FALSE, alternative = "greater") #p=0.02, greater = 0.01, less=0.98
t.test(potential.two, tcga.two, var.equal = FALSE, alternative = "greater") #p=0.0095, greater = 0.0095, less=0.99
```


#weird
weird that the known fusions - one and two vs the sum isn't significantly different
```{r}
t.test(gtex.one, gtex.sum, var.equal = FALSE) #very signif 2.2e-16
t.test(tcga.one, tcga.sum, var.equal = FALSE) #same

t.test(known.one, known.sum, var.equal = FALSE) #p=0.194 
t.test(potential.one, potential.sum, var.equal = FALSE) #p=0.02

t.test(gtex.two, gtex.sum, var.equal = FALSE) #very signif
t.test(tcga.two, tcga.sum, var.equal = FALSE) #same

t.test(known.two, known.sum, var.equal = FALSE) #p=0.194
t.test(potential.two, potential.sum, var.equal = FALSE) #p=0.03

t.test(gtex.one, gtex.two, var.equal = FALSE, alternative = "less") #2.629e-07, greater=1, less=1.315e-07
t.test(tcga.one, tcga.two, var.equal = FALSE, alternative = "less") #0.0189, greater =0.99, less=0.009
t.test(known.one, known.two, var.equal = FALSE, alternative = "less") #0.22, greater =0.88, less=0.11
t.test(potential.one, potential.two, var.equal = FALSE, alternative = "less") #0.022, greater =0.989, less=0.011
```


```{r}
sd(potential.one)
shapiro.test(potential.one) #normal
shapiro.test(log10(potential.one))
shapiro.test(sqrt(potential.one))
shapiro.test(tcga.one)
shapiro.test(sqrt(tcga.one))
shapiro.test(asin(tcga.one))
```


```{r}
ggdensity(potential.one) #not quite bell
ggdensity(log10(potential.one))
ggqqplot(potential.one) #close ish to the line
ggqqplot(log10(potential.one))
ggdensity(tcga.one) #defs not bell
ggdensity(log10(tcga.one))
ggqqplot(tcga.one) #far from line
ggqqplot(log10(tcga.one))
```



#let's try z test for tcga and gtex - because so many samples

```{r}
library(BSDA)
```
it wants populations standard deviation going to have to add summary stats to my stuff
except maybe z test isn't right because I'm not taking a random sample from the population - this is the population?
-or maybe - can do one and two and use total sum sd? - because technically they're samples from the population
-or should I do some ratio stuff?
```{r}
z.test(gtex.sum, tcga.sum, sigma.x = 117.9125, sigma.y = 128.3139)
z.test(tcga.sum, gtex.sum, sigma.x = 128.3139, sigma.y = 117.9125, alternative = "less") #get same p value as done on t test
```

```{r}
sd(gtex.sum) #117.9125
sd(gtex.one) #0.5999231
sd(gtex.two) #0.9500454

sd(tcga.sum) #128.3139
sd(tcga.one) #0.8187869
sd(tcga.two) #1.255165
```


```{r}
#this is gross
#something is wrong because all these columns have different numbers of samples so they shouldn't do that
#oh once it runs out of samples it just repeats it - gross
#don't think I need this I think I can do it without
stats_exon_sums_df <- data.frame(excel_known.df$total_counts, potentialbb.df$total_counts, tcga_ovary_df$total_counts, gtex_ovary_norm_df$total_counts)
colnames(stats_exon_sums_df) <- c("known_fusions", "potential_fusions", "tcga", "gtex")
stats_exon_sums_df
```



```{r}
treatment.na_df <- tcga_ovary_df[is.na(tcga_ovary_df$treatment),]
treated_df <- tcga_ovary_df[!is.na(tcga_ovary_df$treatment),]

t.test(potential.one, treatment.na_df$one, var.equal = FALSE, alternative = "less") #p=0.29, greater=0.15, less=0.85

t.test(treated_df$one, treatment.na_df$one, var.equal = FALSE, alternative = "less") #p=0.38, greater = 0.81, less=0.19
```

looks like google recommends mann-whitney u test - does not assume normality (is a non-parametric test)
http://www.r-tutor.com/elementary-statistics/non-parametric-methods/mann-whitney-wilcoxon-test

```{r}
wilcox.test(potential.one, gtex.one, alternative = "less") #p=0.1015, greater=0.05077, less=0.9507
wilcox.test(potential.one, tcga.one, alternative = "less") #p=0.0551, greater =0.02755, less=0.9727
wilcox.test(tcga.one, gtex.one, alternative = "less") #p=0.0573, greater=0.9714, less=0.02865
wilcox.test(known.one, gtex.one, alternative = "less") #p=0.3105, greater=0.1552, less=0.8492
wilcox.test(known.one, tcga.one, alternative = "less") #p=0.1686, greater=0.08428, less=0.9165
wilcox.test(potential.one, known.one) #p=0.5714 - good
```
-interesting - so tcga.one < gtex.one is stat signif
stat signif as well is that potential.one > tcga.one
-this I didn't expect
and close to being signif is potential.one > gtex.one

```{r}
wilcox.test(potential.sum, known.sum) #p=0.1429 -ooh close but still ok -non-signif - but comparing these 2 I think a t test is fine
wilcox.test(known.sum, gtex.sum, alternative = "greater") #p=0.003316 greater=0.001658
wilcox.test(potential.sum, gtex.sum, alternative = "greater") #p=0.0008688, greater=0.0004344
wilcox.test(known.sum, tcga.sum, alternative = "greater") #p=0.002935, greater=0.001467
wilcox.test(potential.sum, tcga.sum, alternative = "greater") #p=0.0001514, greater=7.569e-05
```

```{r}
wilcox.test(tcga.two, gtex.two, alternative = "less") #p=7.475e-10, greater=1, less=3.737e-10
wilcox.test(potential.two, gtex.two, alternative = "less") #p=0.0002214, greater=0.0001107, less=0.99
wilcox.test(potential.two, tcga.two, alternative = "greater") #p=0.0001294, greater=6.469e-05
wilcox.test(potential.two, known.two, alternative = "greater") #p=1, greater=0.5
wilcox.test(known.two, gtex.two, alternative = "greater") #greater=0.002773
wilcox.test(known.two, tcga.two, alternative = "greater") #greater=0.001739
```
-interesting - seems like exon 2 is really where the change happens

```{r}
wilcox.test(treated_df$total_counts, treatment.na_df$total_counts, alternative = "less") #p=0.5216, greater=0.7396, less=0.2608
wilcox.test(potential.one, treated_df$one, alternative = "less") #p=0.049, greater=0.02454, less=0.9757
wilcox.test(potential.one, treatment.na_df$one, alternative = "less") #warning - can't compute with ties p=0.1763, greater=0.08816, less=0.9177
wilcox.test(treated_df$one, treatment.na_df$one, alternative = "less") #p=0.1578, greater=0.9213, less=0.07891
wilcox.test(treated_df$one, gtex.one, alternative = "less") #p=0.03512, greater=0.9825, less=0.01756
wilcox.test(treatment.na_df$one, gtex.one, alternative = "less") #p=0.8947, greater=0.4474, less=0.5544
wilcox.test(treated_df$total_counts, gtex.sum, alternative = "less") #p<2.2e-16, greater=1, less<2.2e-16
wilcox.test(treatment.na_df$total_counts, gtex.sum, alternative = "less") #p=1.193e-11, greater=1, less=5.967e-12
```


#summary: tests that came up with stat signif results / showed things I wanted
-need to keep in mind that not everything that's statistically significant is biologically significant - and by using non-parametric tests it's not the best measure
-all comparisons potential and known (double sided) non significant - this was as expected
-for others - interested in directional tests: want to know which is larger - normal, cancer, known, potential
```{r}
#TCGA AND GTEX SIGNIF RESULTS
t.test(tcga.sum, gtex.sum, var.equal = FALSE, alternative = "less") #less p<2.2e-16
t.test(tcga.two, gtex.two, var.equal = FALSE, alternative = "less") #less p=9.59e-05
```
-interesting: comparing exon 1 between tcga and gtex - not significant 
-comparing exon 2: tcga < gtex to very high significance

```{r}
t.test(potential.one, gtex.one, var.equal = FALSE, alternative = "greater") #p=0.2, greater=0.1134
t.test(potential.one, tcga.one, var.equal = FALSE, alternative = "less") #p=0.2, greater =0.11
t.test(potential.one[1:2], gtex.one, var.equal = FALSE, alternative = "greater") #very much non-signif
t.test(potential.one[1:2], tcga.one, var.equal = FALSE) #p=0.7373, greater=0.6375, less=0.3687
potential.one[1:2]
potentialbb.df
```
-exon 1 expression is almost significant - interesting because it's greater - could be because only a couple of samples? 
-could try it again with only the 2 most interesting samples - see if it changes - actually got less significant - funny though because the p values flipped around
-maybe a z test would be more appropriate here? - especially with tcga because it's samples out of that population?
-except we can't do a z test because the data isn't normal - despite central limit theorem

-perhaps exon 1 change is significant when comparing the potential to untreated tcga? - that would be awesome


```{r}
treatment.na_df <- tcga_ovary_df[is.na(tcga_ovary_df$treatment),]
treated_df <- tcga_ovary_df[!is.na(tcga_ovary_df$treatment),]

t.test(potential.one, treatment.na_df$one, var.equal = FALSE, alternative = "less") #p=0.29, greater=0.15, less=0.85

t.test(treated_df$one, treatment.na_df$one, var.equal = FALSE, alternative = "less") #p=0.38, greater = 0.81, less=0.19
```

looks like google recommends mann-whitney u test - does not assume normality (is a non-parametric test)
http://www.r-tutor.com/elementary-statistics/non-parametric-methods/mann-whitney-wilcoxon-test

```{r}
wilcox.test(potential.one, gtex.one, alternative = "less") #p=0.1015, greater=0.05077, less=0.9507
wilcox.test(potential.one, tcga.one, alternative = "less") #p=0.0551, greater =0.02755, less=0.9727
wilcox.test(tcga.one, gtex.one, alternative = "less") #p=0.0573, greater=0.9714, less=0.02865
wilcox.test(known.one, gtex.one, alternative = "less") #p=0.3105, greater=0.1552, less=0.8492
wilcox.test(known.one, tcga.one, alternative = "less") #p=0.1686, greater=0.08428, less=0.9165
wilcox.test(potential.one, known.one) #p=0.5714 - good
```
-interesting - so tcga.one < gtex.one is stat signif
stat signif as well is that potential.one > tcga.one
-this I didn't expect
and close to being signif is potential.one > gtex.one

```{r}
wilcox.test(potential.sum, known.sum) #p=0.1429 -ooh close but still ok -non-signif - but comparing these 2 I think a t test is fine
wilcox.test(known.sum, gtex.sum, alternative = "greater") #p=0.003316 greater=0.001658
wilcox.test(potential.sum, gtex.sum, alternative = "greater") #p=0.0008688, greater=0.0004344
wilcox.test(known.sum, tcga.sum, alternative = "greater") #p=0.002935, greater=0.001467
wilcox.test(potential.sum, tcga.sum, alternative = "greater") #p=0.0001514, greater=7.569e-05
```

```{r}
wilcox.test(tcga.two, gtex.two, alternative = "less") #p=7.475e-10, greater=1, less=3.737e-10
wilcox.test(potential.two, gtex.two, alternative = "less") #p=0.0002214, greater=0.0001107, less=0.99
wilcox.test(potential.two, tcga.two, alternative = "greater") #p=0.0001294, greater=6.469e-05
wilcox.test(potential.two, known.two, alternative = "greater") #p=1, greater=0.5
wilcox.test(known.two, gtex.two, alternative = "greater") #greater=0.002773
wilcox.test(known.two, tcga.two, alternative = "greater") #greater=0.001739
```
-interesting - seems like exon 2 is really where the change happens
tcga < gtex - stat signif
potential > tcga and gtex
more signif with tcga - makes sense if tcga < gtex - bigger move
interesting that potential = known - p is 1 and one sided is 0.5

```{r}
wilcox.test(treated_df$total_counts, treatment.na_df$total_counts, alternative = "less") #p=0.5216, greater=0.7396, less=0.2608
wilcox.test(potential.one, treated_df$one, alternative = "less") #p=0.049, greater=0.02454, less=0.9757
wilcox.test(potential.one, treatment.na_df$one, alternative = "less") #warning - can't compute with ties p=0.1763, greater=0.08816, less=0.9177
wilcox.test(treated_df$one, treatment.na_df$one, alternative = "less") #p=0.1578, greater=0.9213, less=0.07891
wilcox.test(treated_df$one, gtex.one, alternative = "less") #p=0.03512, greater=0.9825, less=0.01756
wilcox.test(treatment.na_df$one, gtex.one, alternative = "less") #p=0.8947, greater=0.4474, less=0.5544
wilcox.test(treated_df$total_counts, gtex.sum, alternative = "less") #p<2.2e-16, greater=1, less<2.2e-16
wilcox.test(treatment.na_df$total_counts, gtex.sum, alternative = "less") #p=1.193e-11, greater=1, less=5.967e-12
```

potential one > treated one - signif
treated one < gtex one - signif
(un treated and gtex not signif)
treated total < gtex total - signif



```{r}
tcga_normalised_df
combined_df
```







#4/6

```{r}
hist(tcga_ovary_df$total_counts)
hist(gtex_ovary_df$total_counts)
```

```{r}
tcga_ovary_df %>%
  ggplot(aes(total_counts)) +
  geom_histogram() +
  geom_vline(xintercept = mean(tcga_ovary_df$total_counts),        # Add line for mean
             col = "red",
             lwd = 1) +
geom_vline(xintercept = median(tcga_ovary_df$total_counts),      # Add line for median
             col = "blue",
             lwd = 1)

gtex_ovary_df %>%
  ggplot(aes(total_counts)) +
  geom_histogram() +
  geom_vline(xintercept = mean(gtex_ovary_df$total_counts),        # Add line for mean
             col = "red",
             lwd = 1) +
geom_vline(xintercept = median(gtex_ovary_df$total_counts),      # Add line for median
             col = "blue",
             lwd = 1)
```

so not much of a difference between mean and median
shows outliers nicely though

just want to find a statistical test that will tell me if there's a correlation between low exon 1 and high total expression

```{r}
#mu and sigma here is the tcga ovary mean and sd
#z score - how far is observation from mean
#positive z means greater
#technically data is non normal so z not really appropriate, but have more than enough samples - though group is less so again not quite right test
z.test(tcga_ovary_df[c(1:3),31], mu = 0.247, sigma.x = 0.427) #p <2.2e-16, z = 18.241

z.test(tcga_ovary_df[c(1:3),1], mu = 0.0014, sigma.x = 0.0023) # p= 0.004775, z = 2.8218
sd(tcga_ovary_df[,1])
```




```{r}
sd(tcga_ovary_df[,31])
```

regression maybe? - because we think exon 1 expression affects total?
total is dependent
exon 1 is independent

```{r}
scatter.smooth(x=tcga_ovary_df$one, y=tcga_ovary_df$total_counts, main="Total ~ One")
scatter.smooth(x=gtex_ovary_df$one, y=gtex_ovary_df$total_counts, main="Total ~ One")

par(mfrow=c(1, 2))  # divide graph area in 2 columns
boxplot(tcga_ovary_df$one, main="One", sub=paste("Outlier rows: ", boxplot.stats(tcga_ovary_df$one)$out))  # box plot for 'speed'
boxplot(tcga_ovary_df$total_counts, main="Total", sub=paste("Outlier rows: ", boxplot.stats(tcga_ovary_df$total_counts)$out))  # box plot for 'distance'
```

https://rcompanion.org/handbook/F_12.html

```{r}
if(!require(psych)){install.packages("psych")}
if(!require(mblm)){install.packages("mblm")}
if(!require(quantreg)){install.packages("quantreg")}
if(!require(rcompanion)){install.packages("rcompanion")}
if(!require(mgcv)){install.packages("mgcv")}
if(!require(lmtest)){install.packages("lmtest")}
if(!require(Rfit)){install.packages("Rfit")}
```

```{r}
library(psych)
library(mblm)
library(quantreg)
library(rcompanion)
library(mgcv)
library(lmtest)
library(Rfit)
```

```{r}
library(mblm)
```


```{r}
model.k = mblm(total_counts ~ one,
               data=tcga_ovary_df)

summary(model.k)

model.g = mblm(total_counts ~ one,
               data=gtex_ovary_df)

summary(model.g)
```

#quantile regression
oh didn't realise I've done that before
```{r}
library(quantreg)

model.q = rq(total_counts ~ one,
             data = tcga_ovary_df,
             tau = 0.9) #tau = 0.5 - median

summary(model.q)

#to get p value
model.null = rq(total_counts ~ 1,
                data = tcga_ovary_df,
                tau = 0.9)

anova(model.q, model.null)

#get pseudo r squared
library(rcompanion)

nagelkerke(model.q)

#don't fully understand the quantile regression stuff but I think I like it?? - because I am more interested in the expressed stuff
```


```{r}
#plot
plot(total_counts ~ one,
     data = tcga_ovary_df,
     pch  = 16)

abline(model.q,
       col="blue",
       lwd=2)

library(rcompanion)

Pvalue = anova(model.q, model.null)[[1]][1,4]
Intercept = as.numeric(summary(model.q)$coefficients[1,1])
Slope     = as.numeric(summary(model.q)$coefficients[2,1])
R2     = nagelkerke(model.q)[[2]][3,1]

t1     = paste0("p-value: ", signif(Pvalue, digits=3))
t2     = paste0("R-squared: ", signif(R2, digits=3))
t3     = paste0("Intercept: ", signif(coefficients(model.q)[1], digits=3))
t4     = paste0("Slope: ", signif(coefficients(model.q)[2], digits=3))

text(1160, 2600, labels = t1, pos=4)
text(1160, 2500, labels = t2, pos=4)
text(1160, 2400, labels = t3, pos=4)
text(1160, 2300, labels = t4, pos=4)
```

ahhhh i don't know what I'm doing with these stats
because it's the outliers that are interesting
I suppose this graph is good because it shows super clearly the general trend and the outliers
need to explain to Alicia why I like one on x and total on y - because idea is that exon 1 expression affects total - and explanatory goes on x response on y
maybe if I can add another thing of treated or not treated - multivariable see what that does
but yeah linear regression I don't think really works
and yes z score did show difference so I think can use that and outlier stuff to show it's different



https://statsandr.com/blog/outliers-detection-in-r/#statistical-tests
#there are outlier tests
once again they assume normality
but also screw that I got lots of samples
```{r}
install.packages("outliers")
library(outliers)
```

Grubbs test: - detect if highest or lowest in set is an outlier, default performed on highest
H0: highest is not an outlier
H1: highest is an outlier

```{r}
grubbs.test(tcga_ovary_df$total_counts) #yep outlier
grubbs.test(tcga_ovary_df$one, opposite = TRUE) #highest one is outlier, lowest is not outlier
```

Dixons test - best for small sample sizes < 25

Rosner's test: can detect multiple outliers at once, best for large datasets
```{r}
install.packages("EnvStats")
library(EnvStats)
```

```{r}
rosner <- rosnerTest(tcga_ovary_df$total_counts, k = 10)
rosner$all.stats

#so thought this test was good but actually it says the top 6 results are outliers - I think it's the log transformation that really shows the main ones
```



