---
title: "R Notebook"
output: html_notebook
---

##1 vs total

-TCGA
```{r}
tcga_normalised_df %>%
  ggplot(aes(x = log2(one), y = log2(total_counts), colour = tumour_location)) +
  geom_point() +
  labs(title = "TCGA ABCB1 expression: exon 1 to total counts, organised by tumour")

tcga_normalised_df %>%
  ggplot(aes(x = one, y = log2(total_counts), colour = tumour_location)) +
  geom_point()

tcga_normalised_df %>%
  ggplot(aes(x = log2(one), y = log2(total_counts), colour = one_less_two)) +
  geom_point() +
  facet_wrap(~ tumour_location)
```

-GTEX

```{r}
gtex_normalised_df %>%
  ggplot(aes(x = log2(one), y = log2(total_counts), colour = tissue)) +
  geom_point()

gtex_normalised_df %>%
  ggplot(aes(x = one, y = log2(total_counts), colour = tissue)) +
  geom_point()

gtex_normalised_df %>%
  ggplot(aes(x = log2(one), y = log2(total_counts), colour = one_less_two)) +
  geom_point() +
  facet_wrap(~ tissue)
```


#Boxplots
-Total expression across cancers
```{r}
tcga_normalised_df %>%
  ggplot(aes(x = tumour_location, y = total_counts, colour = tumour_location)) +
  geom_boxplot() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_continuous(breaks = c(0,25,50,75,100,150,200), limits = c(0,225))

gtex_normalised_df %>%
  ggplot(aes(x = tissue, y = total_counts, colour = tissue)) +
  geom_boxplot() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_continuous(breaks = c(0,25,50,75,100,125,150,175,200,225), limits = c(0,225)) +
  labs(title = "GTEx Total counts across tissues")

rbind(tcga_normalised_df[,-c(30,32:33,35:40)], gtex_normalised_df[,-c(30,32:34,36:44)]) %>%
  ggplot(aes(x = tissue, y = total_counts, colour = source)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
tcga_normalised_df$source <- "TCGA"
gtex_normalised_df$source <- "GTEX"
tcga_normalised_df$treated <- !is.na(tcga_normalised_df$treatment)
gtex_normalised_df$treated <- F
gtex_normalised_df
tcga_normalised_df
```


```{r}
combined_df <- rbind(tcga_normalised_df[,-c(30,32:33,35:39)], gtex_normalised_df[,-c(30,32,34,36:44,46)])
combined_df

```


```{r}
combined_df %>%
  filter(tissue != "NA") %>%
  filter(tissue != "Adipose Tissue") %>%
  filter(tissue != "Bile Duct") %>%
  filter(tissue != "Blood Vessel") %>%
  filter(tissue != "Bone Marrow") %>%
  filter(tissue != "Eye") %>%
  filter(tissue != "Head and Neck") %>%
  filter(tissue != "Heart") %>%
  filter(tissue != "Lymph Nodes") %>%
  filter(tissue != "Mesenchymal") %>%
  filter(tissue != "Mesothelium") %>%
  filter(tissue != "Muscle") %>%
  filter(tissue != "Salivary Gland") %>%
  filter(tissue != "Spleen") %>%
  filter(tissue != "Thymus") %>%
  filter(tissue != "") %>%
  filter(tissue != "Vagina") %>%
  ggplot(aes(x = tissue, y = total_counts, colour = source)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_continuous(breaks = c(0,25,50,75,100,125,150,175,200,225), limits = c(0,225)) 

combined_df %>%
  filter(tissue != "NA") %>%
  filter(tissue != "Adipose Tissue") %>%
  filter(tissue != "Bile Duct") %>%
  filter(tissue != "Blood Vessel") %>%
  filter(tissue != "Bone Marrow") %>%
  filter(tissue != "Eye") %>%
  filter(tissue != "Head and Neck") %>%
  filter(tissue != "Heart") %>%
  filter(tissue != "Lymph Nodes") %>%
  filter(tissue != "Mesenchymal") %>%
  filter(tissue != "Mesothelium") %>%
  filter(tissue != "Muscle") %>%
  filter(tissue != "Salivary Gland") %>%
  filter(tissue != "Spleen") %>%
  filter(tissue != "Thymus") %>%
  filter(tissue != "") %>%
  filter(tissue != "Vagina") %>%
  filter(one == 0) %>%
  ggplot(aes(x = tissue, y = total_counts, colour = source)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_continuous(breaks = c(0,25,50,75,100,125,150,175,200,225), limits = c(0,225)) 
#see if I can figure out how to spread it out without losing all the zero's

combined_df %>%
  filter(tissue != "NA") %>%
  filter(tissue != "Adipose Tissue") %>%
  filter(tissue != "Bile Duct") %>%
  filter(tissue != "Blood Vessel") %>%
  filter(tissue != "Bone Marrow") %>%
  filter(tissue != "Eye") %>%
  filter(tissue != "Head and Neck") %>%
  filter(tissue != "Heart") %>%
  filter(tissue != "Lymph Nodes") %>%
  filter(tissue != "Mesenchymal") %>%
  filter(tissue != "Mesothelium") %>%
  filter(tissue != "Muscle") %>%
  filter(tissue != "Salivary Gland") %>%
  filter(tissue != "Spleen") %>%
  filter(tissue != "Thymus") %>%
  filter(tissue != "") %>%
  filter(tissue != "Vagina") %>%
  filter(one > 0) %>%
  ggplot(aes(x = tissue, y = total_counts, colour = source)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_continuous(breaks = c(0,25,50,75,100,125,150,175,200,225), limits = c(0,225)) 

combined_df %>%
  filter(tissue != "NA") %>%
  filter(tissue != "Adipose Tissue") %>%
  filter(tissue != "Bile Duct") %>%
  filter(tissue != "Blood Vessel") %>%
  filter(tissue != "Bone Marrow") %>%
  filter(tissue != "Eye") %>%
  filter(tissue != "Head and Neck") %>%
  filter(tissue != "Heart") %>%
  filter(tissue != "Lymph Nodes") %>%
  filter(tissue != "Mesenchymal") %>%
  filter(tissue != "Mesothelium") %>%
  filter(tissue != "Muscle") %>%
  filter(tissue != "Salivary Gland") %>%
  filter(tissue != "Spleen") %>%
  filter(tissue != "Thymus") %>%
  filter(tissue != "") %>%
  filter(tissue != "Vagina") %>%
  ggplot(aes(x = tissue, y = one, colour = source)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

```{r}
combined_df %>%
  filter(tissue != "NA") %>%
  filter(tissue != "Adipose Tissue") %>%
  filter(tissue != "Bile Duct") %>%
  filter(tissue != "Blood Vessel") %>%
  filter(tissue != "Bone Marrow") %>%
  filter(tissue != "Eye") %>%
  filter(tissue != "Head and Neck") %>%
  filter(tissue != "Heart") %>%
  filter(tissue != "Lymph Nodes") %>%
  filter(tissue != "Mesenchymal") %>%
  filter(tissue != "Mesothelium") %>%
  filter(tissue != "Muscle") %>%
  filter(tissue != "Salivary Gland") %>%
  filter(tissue != "Spleen") %>%
  filter(tissue != "Thymus") %>%
  filter(tissue != "") %>%
  filter(tissue != "Vagina") %>%
  filter(one == 0) %>%
  ggplot(aes(x = tissue, y = total_counts, colour = source)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_continuous(breaks = c(0,25,50,75,100,125,150,175,200,225), limits = c(0,225)) 

combined_df %>%
  filter(tissue != "NA") %>%
  filter(tissue != "Adipose Tissue") %>%
  filter(tissue != "Bile Duct") %>%
  filter(tissue != "Blood Vessel") %>%
  filter(tissue != "Bone Marrow") %>%
  filter(tissue != "Eye") %>%
  filter(tissue != "Head and Neck") %>%
  filter(tissue != "Heart") %>%
  filter(tissue != "Lymph Nodes") %>%
  filter(tissue != "Mesenchymal") %>%
  filter(tissue != "Mesothelium") %>%
  filter(tissue != "Muscle") %>%
  filter(tissue != "Salivary Gland") %>%
  filter(tissue != "Spleen") %>%
  filter(tissue != "Thymus") %>%
  filter(tissue != "") %>%
  filter(tissue != "Vagina") %>%
  filter(one > 0) %>%
  ggplot(aes(x = tissue, y = total_counts, colour = source)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_continuous(breaks = c(0,25,50,75,100,125,150,175,200,225), limits = c(0,225)) 

combined_df %>%
  filter(tissue != "NA") %>%
  filter(tissue != "Adipose Tissue") %>%
  filter(tissue != "Bile Duct") %>%
  filter(tissue != "Blood Vessel") %>%
  filter(tissue != "Bone Marrow") %>%
  filter(tissue != "Eye") %>%
  filter(tissue != "Head and Neck") %>%
  filter(tissue != "Heart") %>%
  filter(tissue != "Lymph Nodes") %>%
  filter(tissue != "Mesenchymal") %>%
  filter(tissue != "Mesothelium") %>%
  filter(tissue != "Muscle") %>%
  filter(tissue != "Salivary Gland") %>%
  filter(tissue != "Spleen") %>%
  filter(tissue != "Thymus") %>%
  filter(tissue != "") %>%
  filter(tissue != "Vagina") %>%
  filter(source == "TCGA") %>%
  ggplot(aes(x = tissue, y = total_counts, colour = one_not_expressed)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_continuous(breaks = c(0,25,50,75,100,125,150,175,200,225), limits = c(0,225)) 

#can see a bit of a trend
#tissues where it's normally expressed in GTEX (adrenal gland, liver, colon, intestine all that stuff) - see higher expression when 1=0
#tissues where it's not normally expressed - see more of exon 1 expressed
#similar trend in cancer graph above - where we know cancer is overexpressed - see higher
#will need to talk about these
combined_df %>%
  filter(tissue != "NA") %>%
  filter(tissue != "Adipose Tissue") %>%
  filter(tissue != "Bile Duct") %>%
  filter(tissue != "Blood Vessel") %>%
  filter(tissue != "Bone Marrow") %>%
  filter(tissue != "Eye") %>%
  filter(tissue != "Head and Neck") %>%
  filter(tissue != "Heart") %>%
  filter(tissue != "Lymph Nodes") %>%
  filter(tissue != "Mesenchymal") %>%
  filter(tissue != "Mesothelium") %>%
  filter(tissue != "Muscle") %>%
  filter(tissue != "Salivary Gland") %>%
  filter(tissue != "Spleen") %>%
  filter(tissue != "Thymus") %>%
  filter(tissue != "") %>%
  filter(tissue != "Vagina") %>%
  filter(source == "GTEX") %>%
  ggplot(aes(x = tissue, y = total_counts, colour = one_not_expressed)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_continuous(breaks = c(0,25,50,75,100,125,150,175,200,225), limits = c(0,225)) 
```





```{r}
tcga_ovary_df %>%
  ggplot(aes(total_counts)) +
  geom_histogram() +
  geom_vline(xintercept = mean(tcga_ovary_df$total_counts),        # Add line for mean
             col = "red",
             lwd = 1) +
geom_vline(xintercept = median(tcga_ovary_df$total_counts),      # Add line for median
             col = "blue",
             lwd = 1)

combined_df %>%
  filter(tissue != "NA") %>%
  filter(tissue != "Adipose Tissue") %>%
  filter(tissue != "Bile Duct") %>%
  filter(tissue != "Blood Vessel") %>%
  filter(tissue != "Bone Marrow") %>%
  filter(tissue != "Eye") %>%
  filter(tissue != "Head and Neck") %>%
  filter(tissue != "Heart") %>%
  filter(tissue != "Lymph Nodes") %>%
  filter(tissue != "Mesenchymal") %>%
  filter(tissue != "Mesothelium") %>%
  filter(tissue != "Muscle") %>%
  filter(tissue != "Salivary Gland") %>%
  filter(tissue != "Spleen") %>%
  filter(tissue != "Thymus") %>%
  filter(tissue != "") %>%
  filter(tissue != "Vagina") %>%
  ggplot(aes(x = total_counts, colour = source)) +
  geom_histogram(fill = "white", position = "dodge") +
  facet_wrap(~ tissue) 
#didn't quite work
```


-Total and exon1
```{r}
tcga_normalised_df %>%
  ggplot(aes(x = tumour_location, y = total_counts)) +
  geom_boxplot() +
  scale_y_continuous(breaks = c(0,25,50,75,100,150,200), limits = c(0,225))

gtex_normalised_df %>%
  ggplot(aes(x = tissue, y = total_counts)) +
  geom_boxplot() +
  scale_y_continuous(breaks = c(0,25,50,75,100,150,200), limits = c(0,225))

tcga_normalised_df %>%
  ggplot(aes(x = tumour_location, y = one)) +
  geom_boxplot() +
  scale_y_continuous(breaks = c(0, 0.025, 0.05, 0.075, 0.1, 0.15), limits = c(0, 0.175))

gtex_normalised_df %>%
  ggplot(aes(x = tissue, y = one)) +
  geom_boxplot() +
  scale_y_continuous(breaks = c(0, 0.025, 0.05, 0.075, 0.1, 0.15), limits = c(0, 0.175))
```

```{r}
combined_df %>%
  group_by(tissue) %>%
  summarise_at(vars(total_counts))

combined_df %>%
  summarise(mean(total_counts))

summary(combined_df[])
```




##OVARY (+ Fallopian)

```{r}
tcga_ovary_df <- subset(tcga_normalised_df, tumour_location == "Ovary")
gtex_ovary_df <- subset(gtex_normalised_df, tissue == "Ovary")
gtex_fallopian_df <- subset(gtex_normalised_df, tissue == "Fallopian Tube")

tcga_ovary_long <- tcga_ovary_df[-c(30:36)]
tcga_ovary_long <- data.frame(t(tcga_ovary_long))
tcga_ovary_long$exon_number <- c(1:29)

gtex_ovary_long <- gtex_ovary_df[-c(30:42)]
gtex_ovary_long <- data.frame(t(gtex_ovary_long))
gtex_ovary_long$exon_number <- c(1:29)

gtex_fallopian_long <- gtex_fallopian_df[-c(30:42)]
gtex_fallopian_long <- data.frame(t(gtex_fallopian_long))
gtex_fallopian_long$exon_number <- c(1:29)
```

-graphs
```{r}
tcga_ovary_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-16:-2), limits = c(-18,0))

gtex_ovary_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-16:-2), limits = c(-18,0))

gtex_fallopian_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-16:-2), limits = c(-18,0))
```

##PROSTATE (CHIRSTIE IS INTERESTED IN PROSTATE)

```{r}
tcga_prostate_df <- subset(tcga_normalised_df, tissue == "Prostate")
gtex_prostate_df <- subset(gtex_normalised_df, tissue == "Prostate")

tcga_prostate_long <- tcga_prostate_df[-c(30:36)]
tcga_prostate_long <- data.frame(t(tcga_prostate_long))
tcga_prostate_long$exon_number <- c(1:29)

gtex_prostate_long <- gtex_prostate_df[-c(30:42)]
gtex_prostate_long <- data.frame(t(gtex_prostate_long))
gtex_prostate_long$exon_number <- c(1:29)
```

```{r}
summary(tcga_prostate_df$total_counts)
summary(gtex_prostate_df$total_counts)

tcga_prostate_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-15:0), limits = c(-15,0)) +
  labs(title = "TCGA Prostate data")

tcga_prostate_long[,c(1:5,559)] %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-15:0), limits = c(-15,0)) +
  labs(title = "TCGA Prostate data - top 5 total expression")

gtex_prostate_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-15:0), limits = c(-15,0)) +
  labs(title = "GTEx Prostate data")
```

##LIVER (PAPER FOUND FUSIONS IN LIVER)

```{r}
tcga_liver_df <- subset(tcga_normalised_df, tissue == "Liver")
gtex_liver_df <- subset(gtex_normalised_df, tissue == "Liver")

tcga_liver_long <- tcga_liver_df[-c(30:36)]
tcga_liver_long <- data.frame(t(tcga_liver_long))
tcga_liver_long$exon_number <- c(1:29)

gtex_liver_long <- gtex_liver_df[-c(30:42)]
gtex_liver_long <- data.frame(t(gtex_liver_long))
gtex_liver_long$exon_number <- c(1:29)
```

```{r}
tcga_liver_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-15:2), limits = c(-16,3)) +
  labs(title = "TCGA Liver data")

gtex_liver_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-15:2), limits = c(-16,3)) +
  labs(title = "GTEx Liver data")
```
##BLOOD (ALICIA IS INTERESTED IN BLOOD)

```{r}
distinct(tcga_normalised_df, tumour_location)
distinct(gtex_normalised_df, tissue)
```


```{r}
tcga_blood_df <- subset(tcga_normalised_df, tissue == "Blood")
gtex_blood_df <- subset(gtex_normalised_df, tissue == "Blood")

tcga_blood_long <- tcga_blood_df[-c(30:36)]
tcga_blood_long <- data.frame(t(tcga_blood_long))
tcga_blood_long$exon_number <- c(1:29)

gtex_blood_long <- gtex_blood_df[-c(30:42)]
gtex_blood_long <- data.frame(t(gtex_blood_long))
gtex_blood_long$exon_number <- c(1:29)
```

```{r}
tcga_blood_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-16:0), limits = c(-17,1)) +
  labs(title = "TCGA Blood data")

gtex_blood_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-16:0), limits = c(-17,1)) +
  labs(title = "GTEx Blood data")
```
##BREAST (CHRISTIE ALSO FOUND FUSIONS IN BREAST)

```{r}
tcga_breast_df <- subset(tcga_normalised_df, tissue == "Breast")
gtex_breast_df <- subset(gtex_normalised_df, tissue == "Breast")

tcga_breast_long <- tcga_breast_df[-c(30:36)]
tcga_breast_long <- data.frame(t(tcga_breast_long))
tcga_breast_long$exon_number <- c(1:29)

gtex_breast_long <- gtex_breast_df[-c(30:42)]
gtex_breast_long <- data.frame(t(gtex_breast_long))
gtex_breast_long$exon_number <- c(1:29)
```

```{r}
tcga_breast_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-16:0), limits = c(-16,1)) +
  labs(title = "TCGA Breast data")

gtex_breast_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-16:0), limits = c(-16,1)) +
  labs(title = "GTEx Breast data")
```
##BLADDER (from my lit review list)

```{r}
tcga_bladder_df <- subset(tcga_normalised_df, tissue == "Bladder")
gtex_bladder_df <- subset(gtex_normalised_df, tissue == "Bladder")

tcga_bladder_long <- tcga_bladder_df[-c(30:36)]
tcga_bladder_long <- data.frame(t(tcga_bladder_long))
tcga_bladder_long$exon_number <- c(1:29)

gtex_bladder_long <- gtex_bladder_df[-c(30:42)]
gtex_bladder_long <- data.frame(t(gtex_bladder_long))
gtex_bladder_long$exon_number <- c(1:29)
```

```{r}
tcga_bladder_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-15:-1), limits = c(-15,-1)) +
  labs(title = "TCGA Bladder data")

gtex_bladder_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-15:-1), limits = c(-15,-1)) +
  labs(title = "GTEx Bladder data")
```
```{r}
tcga_bladder_df
#top sample rail_66690 TCGA:TCGA-BT-A20W-11A no treatment data
#total = 7.34, one = 0.018
```

##TESTIS (lit review)

-hello, this is the first clear outlier since Ovary
```{r}
tcga_testis_df <- subset(tcga_normalised_df, tissue == "Testis")
gtex_testis_df <- subset(gtex_normalised_df, tissue == "Testis")

tcga_testis_long <- tcga_testis_df[-c(30:36)]
tcga_testis_long <- data.frame(t(tcga_testis_long))
tcga_testis_long$exon_number <- c(1:29)

gtex_testis_long <- gtex_testis_df[-c(30:42)]
gtex_testis_long <- data.frame(t(gtex_testis_long))
gtex_testis_long$exon_number <- c(1:29)
```

```{r}
tcga_testis_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-14:1), limits = c(-15,1)) +
  labs(title = "TCGA Testis data")

gtex_testis_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-14:1), limits = c(-15,1)) +
  labs(title = "GTEx Testis data")
```
```{r}
tcga_testis_df
#rail_67633 TCGA: TCGA-2G-AAG6-01A no treatment data
#one = 0, total = 33.93
#rail_63532 TCGA: TCGA-2G-AALO-01A no treatment data
#one = 0, total = 12.58
#3rd highest has total of 6
```


##LUNG - lit review
-another nice outlier sitting on top of the rest
```{r}
tcga_lung_df <- subset(tcga_normalised_df, tissue == "Lung")
gtex_lung_df <- subset(gtex_normalised_df, tissue == "Lung")

tcga_lung_long <- tcga_lung_df[-c(30:36)]
tcga_lung_long <- data.frame(t(tcga_lung_long))
tcga_lung_long$exon_number <- c(1:29)

gtex_lung_long <- gtex_lung_df[-c(30:42)]
gtex_lung_long <- data.frame(t(gtex_lung_long))
gtex_lung_long$exon_number <- c(1:29)
```

```{r}
tcga_lung_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-15:0), limits = c(-15,1)) +
  labs(title = "TCGA Lung data")

gtex_lung_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-15:0), limits = c(-15,1)) +
  labs(title = "GTEx Lung data")
```
```{r}
tcga_lung_df
#rail_61921, TCGA: TCGA-50-5066-02A, treatment = Avastin
#one = 0, total = 18.78
#next highest 11.3 - but less obviously outlier
```


##GASTRIC - from lit review (Stomach)

-wow massively obvious difference between gtex and tcga - lots of tcga much higher
-also some nice clear outliers from tcga

```{r}
tcga_stomach_df <- subset(tcga_normalised_df, tissue == "Stomach")
gtex_stomach_df <- subset(gtex_normalised_df, tissue == "Stomach")

tcga_stomach_long <- tcga_stomach_df[-c(30:36)]
tcga_stomach_long <- data.frame(t(tcga_stomach_long))
tcga_stomach_long$exon_number <- c(1:29)

gtex_stomach_long <- gtex_stomach_df[-c(30:42)]
gtex_stomach_long <- data.frame(t(gtex_stomach_long))
gtex_stomach_long$exon_number <- c(1:29)
```

```{r}
tcga_stomach_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-16:1), limits = c(-17,1)) +
  labs(title = "TCGA Stomach data")

gtex_stomach_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-16:1), limits = c(-17,1)) +
  labs(title = "GTEx Stomach data")
```
```{r}
tcga_stomach_df
#basically all the red are outliers - if compare to GTEx
#rail_64862 TCGA: TCGA-BR-8682-01A treatment = Xeloda
#one = 0.024, total = 45.46
#rail_64380 TCGA: TCGA-CG-4436-01A treatment = NA
# one = 0.0014, total = 31.67
#2 in 20's, 4 in 10s
```


##CERVICAL - lit review
-yay got some outliers once again
```{r}
tcga_cervix_df <- subset(tcga_normalised_df, tissue == "Cervix")
gtex_cervix_df <- subset(gtex_normalised_df, tissue == "Cervix Uteri")

tcga_cervix_long <- tcga_cervix_df[-c(30:36)]
tcga_cervix_long <- data.frame(t(tcga_cervix_long))
tcga_cervix_long$exon_number <- c(1:29)

gtex_cervix_long <- gtex_cervix_df[-c(30:42)]
gtex_cervix_long <- data.frame(t(gtex_cervix_long))
gtex_cervix_long$exon_number <- c(1:29)
```

```{r}
tcga_cervix_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-15:0), limits = c(-15,1)) +
  labs(title = "TCGA Cervix data")

gtex_cervix_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-15:0), limits = c(-15,1)) +
  labs(title = "GTEx Cervix data")
```
```{r}
tcga_cervix_df
#rail_69840 TCGA: TCGA-FU-A3EO-11A no treatment
#one = 0.0084, total = 14.29
#rail_63385 TCGA: TCGA-VS-A9V4-01A treatment = Cisplatin
#one = 0, total = 10.27
#7, 6, 4, then lots of 2s
```


##KIDNEY - lit review (not in table but should have been maybe)
-wow got some extreme overexpression compared to gtex
```{r}
tcga_kidney_df <- subset(tcga_normalised_df, tissue == "Kidney")
gtex_kidney_df <- subset(gtex_normalised_df, tissue == "Kidney")

tcga_kidney_long <- tcga_kidney_df[-c(30:36)]
tcga_kidney_long <- data.frame(t(tcga_kidney_long))
tcga_kidney_long$exon_number <- c(1:29)

gtex_kidney_long <- gtex_kidney_df[-c(30:42)]
gtex_kidney_long <- data.frame(t(gtex_kidney_long))
gtex_kidney_long$exon_number <- c(1:29)
```

```{r}
tcga_kidney_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-13:3), limits = c(-14,4)) +
  labs(title = "TCGA Kidney data")

gtex_kidney_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-13:3), limits = c(-14,4)) +
  labs(title = "GTEx Kidney data")
```



##UTERUS/VAGINA - not in my lists but because it's close to ovary location wise would like to look - TCGA has uterus, gtex has vagina and uterus
-doesn't really look like anything interesting happening
```{r}
tcga_uterus_df <- subset(tcga_normalised_df, tumour_location == "Uterus")
gtex_uterus_df <- subset(gtex_normalised_df, tissue == "Uterus")

tcga_uterus_long <- tcga_uterus_df[-c(30:36)]
tcga_uterus_long <- data.frame(t(tcga_uterus_long))
tcga_uterus_long$exon_number <- c(1:29)

gtex_uterus_long <- gtex_uterus_df[-c(30:42)]
gtex_uterus_long <- data.frame(t(gtex_uterus_long))
gtex_uterus_long$exon_number <- c(1:29)
```

```{r}
tcga_uterus_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-15:1), limits = c(-16,2)) +
  labs(title = "TCGA Uterus data")

gtex_uterus_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-15:1), limits = c(-16,2)) +
  labs(title = "GTEx Uterus data")
```
```{r}
tcga_normalised_df
gtex_normalised_df
```

##Esophagus
```{r}
tcga_esophagus_df <- subset(tcga_normalised_df, tissue == "Esophagus")
gtex_esophagus_df <- subset(gtex_normalised_df, tissue == "Esophagus")

tcga_esophagus_long <- tcga_esophagus_df[,-c(30:40)]
tcga_esophagus_long <- data.frame(t(tcga_esophagus_long))
tcga_esophagus_long$exon_number <- c(1:29)

gtex_esophagus_long <- gtex_esophagus_df[,-c(30:47)]
gtex_esophagus_long <- data.frame(t(gtex_esophagus_long))
gtex_esophagus_long$exon_number <- c(1:29)


tcga_esophagus_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-15:0), limits = c(-15,0)) +
  labs(title = "TCGA Esophagus cancer data")

gtex_salivary_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-15:0), limits = c(-15,0)) +
  labs(title = "GTEx Esophagus data")
```
```{r}
tcga_colon_df <- subset(tcga_normalised_df, tissue == "Colorectal")
gtex_colon_df <- subset(gtex_normalised_df, tissue == "Colon")
```


##Brain
```{r}
tcga_brain_df <- subset(tcga_normalised_df, tissue == "Brain")
gtex_brain_df <- subset(gtex_normalised_df, tissue == "Brain")

tcga_brain_long <- tcga_brain_df[,-c(30:40)]
tcga_brain_long <- data.frame(t(tcga_brain_long))
tcga_brain_long$exon_number <- c(1:29)

gtex_brain_long <- gtex_brain_df[,-c(30:47)]
gtex_brain_long <- data.frame(t(gtex_brain_long))
gtex_brain_long$exon_number <- c(1:29)


tcga_brain_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  labs(title = "TCGA Brain cancer data")

gtex_brain_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  labs(title = "GTEx Brain data")
```
##Pancreas
```{r}
tcga_pancreas_df <- subset(tcga_normalised_df, tissue == "Pancreas")
gtex_pancreas_df <- subset(gtex_normalised_df, tissue == "Pancreas")

tcga_pancreas_long <- tcga_pancreas_df[,-c(30:40)]
tcga_pancreas_long <- data.frame(t(tcga_pancreas_long))
tcga_pancreas_long$exon_number <- c(1:29)

gtex_pancreas_long <- gtex_pancreas_df[,-c(30:47)]
gtex_pancreas_long <- data.frame(t(gtex_pancreas_long))
gtex_pancreas_long$exon_number <- c(1:29)


tcga_pancreas_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  labs(title = "TCGA Pancreas cancer data")

gtex_pancreas_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  labs(title = "GTEx Pancreas data")
```
##Skin
```{r}
tcga_skin_df <- subset(tcga_normalised_df, tissue == "Skin")
gtex_skin_df <- subset(gtex_normalised_df, tissue == "Skin")

tcga_skin_long <- tcga_skin_df[,-c(30:40)]
tcga_skin_long <- data.frame(t(tcga_skin_long))
tcga_skin_long$exon_number <- c(1:29)

gtex_skin_long <- gtex_skin_df[,-c(30:47)]
gtex_skin_long <- data.frame(t(gtex_skin_long))
gtex_skin_long$exon_number <- c(1:29)


tcga_skin_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  labs(title = "TCGA Skin cancer data")

gtex_skin_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  labs(title = "GTEx Skin data")
```





##HEAD AND NECK - paper had a fusion there
-theres esophagus data as well in both which is vaguely in head and neck area
```{r}
tcga_head_df <- subset(tcga_normalised_df, tumour_location == "Head and Neck")
gtex_salivary_df <- subset(gtex_normalised_df, tissue == "Salivary Gland")

tcga_head_long <- tcga_head_df[-c(30:36)]
tcga_head_long <- data.frame(t(tcga_head_long))
tcga_head_long$exon_number <- c(1:29)

gtex_salivary_long <- gtex_salivary_df[-c(30:42)]
gtex_salivary_long <- data.frame(t(gtex_salivary_long))
gtex_salivary_long$exon_number <- c(1:29)
```

```{r}
tcga_head_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-15:0), limits = c(-15,0)) +
  labs(title = "TCGA Head and Neck cancer data")

gtex_salivary_long %>%
  melt(id.vars = "exon_number") %>%
  ggplot(aes(x = exon_number, y = log2(value), group = variable, colour = variable)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(1:29), limits = c(1,29)) +
  scale_y_continuous(breaks = c(-15:0), limits = c(-15,0)) +
  labs(title = "GTEx Salivary Gland data")
```

```{r}
tcga_ovary_df %>%
  ggplot(aes(x = one, y = total_counts)) +
  geom_point()

gtex_ovary_df %>%
  ggplot(aes(x = one, y = total_counts)) +
  geom_point()
```

```{r}
distinct(tcga_normalised_df, tumour_location)
distinct(gtex_normalised_df, tissue)
```

-kidney
-adrenal gland
-liver
-stomach
-testis
-uterus
-lung
-cervix
-breast
-prostate
-bladder
-ovary









